---
title: "Prediction of Activity Type"
output: html_document
---

This report uses data collected by devices detecting human movements. I described how I built my model, how I did cross validation and estimated error, and how I made the choices on these steps. Then I predicted activity types with the testing dataset.

First read data:

```{r,cache=TRUE}
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")
library(caret)
```

Then I processed data. I found there are NA values in the testing set so I excluded those columns from both training and testing data, using near zero variation. In addition, the first column "X" in both datasets are unique ids for each observation. I excluded it from bot datasets as well because it would heavily biase the prediction results.

```{r,cache=TRUE}
#calculating near zero variation (NAs in the testing set):
nzv <- nearZeroVar(testing,saveMetrics=TRUE)
#keeping non NAs in both datasets:
testingnew <- testing[,which(nzv$zeroVar=="FALSE")]
trainingnew <- training[,which(nzv$zeroVar=="FALSE")]
#removing the first column "X" from both datasets:
testingnew <- testingnew[,-1]
trainingnew<-trainingnew[,-1]
```

I sliced training data into training and validation sets:

```{r,cache=TRUE}
intrain <- createDataPartition(y=trainingnew$classe,p=0.9,list=FALSE)
train<-trainingnew[intrain,]
valid<-trainingnew[-intrain,]
```

I chose to fit a random forests model because it is precise and works better with non-linear models. Then I did prediction with the validation data.

```{r,cache=TRUE}
model <- train(classe~.,data=train,method="rf")
validate <- predict(model,newdata=valid)
confusionMatrix(validate,valid$classe)
```

It appears that the prediction is accurate. Then I did cross valdiation with rfcv:

```{r,cache=TRUE}
crossval <- rfcv(trainingnew[,-58],trainingnew[,58],cv.fold=5)
crossval[2]
```

So the expected error rate is around the errors shown above.
Predicting test cases:

```{r,cache=TRUE}
answers <- predict(model,newdata=testingnew)
answers
#Writing answers into text files:
pml_write_files=function(x){
      n=length(x)
      for (i in 1:n){
            filename=paste0("problem_id_",i,".txt")
            write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
     }
}
pml_write_files(answers)
```
